---
title: "hyperSpec Package Review"
author: "Devin Gordon"
date: "Thursday, March 05, 2015"
output: ioslides_presentation
---

<!--
# Script Name: hyperSpecReview.Rmd
# Purpose: This is an R Markdown to review the utilities of the hyperSpec R package
# Authors: Devin A. Gordon
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - Devin Gordon started blank script
# v0.00.02 - Presentation renamed to V2 and changed to html output

# Rmd code goes below the comment marker!
-->

#Background

## Spectra - What are they?

- "a specific set of values that can vary infinitely within a continuum"
- Generally: 2-dimensional (sometimes higher) dimensional data, where the first dimension identifies points along a continuum, and the second dimension the values of a response of interest at the corresponding points
- Examples: Optical, Mass, Political Alignment

## Spectral Data - How do we store them?

- Proprietary file format: not so useful
- .csv files (ASCII): better ~ excel, R
- .spc files (Binary): good ~ R

## hyperSpec - How to manipulate .spc files in R

- http://hyperspec.r-forge.r-project.org/
- Handles spectral data
- Features: Convenient import, spectral range selection, shfiting,  plotting/viewing, normalization, smoothing, correction, arithmetic, initial data analysis


#hyperSpec Examples
```{r, echo=FALSE}
library(hyperSpec)
```

## File Import

```{r}
# Read one .spc file and check its structure
setwd("V:/vuv-data/instr/spectra/")
curfile <- "./Fluorescence/initial_pet_Trial/corrected_pet_unstab_cyclic_quv_ex320/sa19603_00_ex320.spc"
step0 <- read.spc(curfile)
class(step0)
step0
```

## Basic Plotting

```{r}
#Plot with hyperSpec base plotting

plotspc(step0)
```

## Multiple .spc Files?

- (Purposefully without a loop)
```{r}
# Read multiple .spc files

setwd("./data/corrected_pet_unstab_cyclic_quv_ex320/")
curfile <- "./sa19603_00_ex320.spc"
step0 <- read.spc(curfile)
curfile <- "./sa19603_22_ex320.spc"
step1 <- read.spc(curfile)
curfile <- "./sa19603_23_ex320.spc"
step2 <- read.spc(curfile)
curfile <- "./sa19603_24_ex320.spc"
step3 <- read.spc(curfile)
curfile <- "./sa19603_25_ex320.spc"
step4 <- read.spc(curfile)
curfile <- "./sa19603_26_ex320.spc"
step5 <- read.spc(curfile)
curfile <- "./sa19603_27_ex320.spc"
step6 <- read.spc(curfile) 
curfile <- "./sa19603_28_ex320.spc"
step7 <- read.spc(curfile)
```

## Combine into one object (Good News!)

```{r}
# Use Collpase to combine spectra

spec <- collapse(step0, step1, step2, step3, step4, step5, step6, step7)
class(spec)
spec
```

## Basic Plotting

```{r}
plotspc(spec)
```

## Spectral Range Selection

```{r}
#Subset and redifine the hyperSpec object according to wavelength

spec <- spec [,, min ~ 600]
spec
```

## Basic Plotting


```{r}
plotspc(spec)
```

## Spectral Arithmetic (Subtraction)
```{r}
#Perform spectral subtraction

sub <- step7-step0
plotspc(sub[,, min ~ 600])
```

## Removing Bad Data
```{r}
# Let's look at some fake IR data
ir.spc <- chondro
ir.spc
plotspc(ir.spc, spc.nmax = length(ir.spc))
```

## Removing Bad Data
```{r}
#Define any point above 1500 as bad
high.int <- apply (ir.spc > 1500, 1, any)

#Maximum should atleast be 0.1 
low.int <- apply (ir.spc, 1, max) < 0.1

#Apply Conditions
ir.spc <- ir.spc [! high.int & ! low.int]
```

## Removed the bad spectra
```{r}
plotspc(ir.spc, spc.nmax = length(ir.spc))
```

## Baseline Correction (One of many options)
```{r}
# Apply basline correction function
blcorr <- spc.fit.poly.below (ir.spc)

# Subtract away correction from original
ir.spc <- ir.spc - blcorr
```

## Corrected Spectra!
```{r}
plotspc(ir.spc)
```

## Matrix Plot - IR Data
```{r}
plotmat(ir.spc)
```

## Stacked Spectra

```{r}
plotspc(spec, stacked = TRUE)
```

# Summary

## hyperSpec

- hyperSpec is a powerful R package for handling and manipulating spectral data in R
- More to come!

